Neural Models for Target-Based Computer-Assisted Musical Orchestration: A Preliminary Study

### 2020-08-19 paper/aimusic2020/paper.tex
Reduced length of paper

\begin{figure}
\center{
\includegraphics[scale=0.4]{../latex/figs/CNN_vs_ResNet.png}}
\caption{Best overall accuracy for CNN with LSTM and ResNet depending on the number of instruments in the combinations used for training. \label{cnn_vs_resnet}}
\end{figure}

@@

The choice of using the Mel spectrogram and MFCCs as input features for classification models is common in music information retrieval \cite{McKinney2003} and can be considered to be an appropriate representation of sound and musical signals. The Mel spectrogram is generated using an FFT hop length of 2048 samples (the window of each FFT was $46$ms wide), and a number of Mel bins equal to 128. Therefore, the Mel features fed to the model were matrices of size $128\times 345$. We used the Librosa package in Python to compute the features; more details on the exact computations can be found in \cite{mcfee15}. The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms would be allowed. Such setup proved to be successful in \cite{Salamon17}.

@@

Following the first three convolutional layers, there is an LSTM layer which outputs 32 matrices. After the LSTM layer, there is a final convolutional layer yielding a tensor of dimensions $32\times 8 \times 21$. We flatten the outputs and feed them into fully a connected layer with Dropout, then another another fully connected layer. Finally, the sigmoid function is applied to the final layer.

@@

During training, the loss function used to optimize the inner parameters of the model was binary cross entropy, as it is the common choice for multiclass multilabel classification frameworks. However, the value of the loss function alone is difficult to interpret. For this reason we created a complementary function $f$ to be used for evaluation only. This function compares a vector of expected outputs $\overline{X}$ with the estimated output from the model $\hat{X}$ by using the following function

\begin{equation}
f(\overline{X}, \hat{X}) = \frac{1}{N}<\overline{X}, M_N(\hat{X})>
\label{eval}
\end{equation}
\begin{equation}
M_N(\hat{X})_i = \left\{\begin{array}{ll}
1 \text{ if } i \in I_N(\hat{X})\\
0 \text{ otherwise}
\end{array}\right.
\label{NMax}
\end{equation}
and $I_N(\hat{X})$ is the set of indices containing the $N$ first maximums of the vector $\hat{X}$.

More specifically, the function $M_N(\hat{X})$ takes as an input a vector of probabilities and outputs a vector where only the positions of the $N$ first maxima are set to 1. This new vector would be the orchestration of $N$ instruments given by the model. Thus, the function $f$ simply outputs the proportion of the estimated orchestration that matches the expected one.

@@

Then, for both CNN with LSTM and ResNet, we computed the maximum accuracy over the epochs. 
The CNN was trained on 200,000 generated samples over 50 epochs, and ResNet was 	trained on 400,000 generated samples over 20 epochs. Fig. \ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of samples $N$.

ResNet outperforms the CNN regardless of the number of samples used in the combination. This result is consistent with previous research \cite{He15}, as residual networks usually perform well in classification problems.

@@

While it is not easy to explain these differences in accuracy, we hypothesize this being related to the nature of peculiar spectral and temporal morphology of each instrument. For example, flute notes tend to exhibit a steep spectral rolloff, with most of the energy captured by the first few partials. Moreover, the noisy nature of the transient portions of these note is not well represented by frequency-based descriptions such as Mel spectra. These two factors combined, could make the disentanglement of the flute from the analyzed combination more difficult.

@@

Orchidea implements many advanced features that are not supported by our models. For example, it is able to apply symbolic constraints to the search, hence allowing only specific instrumental combinations or playing styles. It is also able to reduce the search space by applying harmonic analysis on the target sound, and create sparse solutions that do not use every instrument specified.

In order to have a fairer comparison, we did not allow Orchidea to use any of the advanced features: we did not apply any symbolic constraints or harmonic analysis and we forced it forced to use all 10 instruments in each solution. This creates a more fair comparison, since our models are unable to create constrained or sparse solutions.

>>>>

@@

The choice of using the Mel spectrogram and MFCCs as input features for classification models is common in music information retrieval \cite{McKinney2003} and can be considered to be an appropriate representation of sound and musical signals. The Mel spectrogram is generated using an FFT hop length of 2048 samples (the window of each FFT was $46$ms wide), and a number of Mel bins equal to 128. Therefore, the Mel features fed to the model were matrices of size $128\times 345$. We used the Librosa package in Python to compute the features; more details on the exact computations can be found in \cite{mcfee15}. %The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms would be allowed.
Such setup proved to be successful in \cite{Salamon17}.

@@

Following the first three convolutional layers, there is an LSTM layer which outputs 32 matrices. After the LSTM layer, there is a final convolutional layer yielding a tensor of dimensions $32\times 8 \times 21$. We flatten the outputs and feed them into a fully connected layer with Dropout, then another another fully connected layer. Finally, the sigmoid function is applied to the final layer.

@@

During training, the loss function used to optimize the inner parameters of the model was binary cross entropy, as it is the common choice for multiclass multilabel classification frameworks. However, the value of the loss function alone is difficult to interpret.
%For this reason we created a complementary function $f$ to be used for evaluation only. This function compares a vector of expected outputs $\overline{X}$ with the estimated output from the model $\hat{X}$ by using the following function
%
%\begin{equation}
%f(\overline{X}, \hat{X}) = \frac{1}{N}<\overline{X}, M_N(\hat{X})>
%\label{eval}
%\end{equation}
%\begin{equation}
%M_N(\hat{X})_i = \left\{\begin{array}{ll}
%1 \text{ if } i \in I_N(\hat{X})\\
%0 \text{ otherwise}
%\end{array}\right.
%\label{NMax}
%\end{equation}
%and $I_N(\hat{X})$ is the set of indices containing the $N$ first maximums of the vector $\hat{X}$. More specifically, the function $M_N(\hat{X})$ takes as an input a vector of probabilities and outputs a vector where only the positions of the $N$ first maxima are set to 1. This new vector would be the orchestration of $N$ instruments given by the model. Thus, the function $f$ simply outputs the proportion of the estimated orchestration that matches the expected one.
For this reason, the evaluation was done by comparing the proportion of estimated orchestration samples, chosen among the samples that output the highest probability, that matched the expected orchestration.

@@

Then, for both CNN with LSTM and ResNet, we computed the maximum accuracy over the epochs. 
%The CNN was trained on 200,000 generated samples over 50 epochs, and ResNet was 	trained on 400,000 generated samples over 20 epochs. Fig. \ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of samples $N$.
ResNet outperforms the CNN regardless of the number of samples used in the combination. This result is consistent with previous research \cite{He15}, as residual networks usually perform well in classification problems.

\begin{figure}
\center{
\includegraphics[scale=0.4]{../latex/figs/CNN_vs_ResNet.png}}
\caption{Best overall accuracy for CNN with LSTM (50 epochs, 200k samples per epoch) and ResNet (20 epochs, 400k samples per epoch) depending on the number of instruments in the combinations used for training. \label{cnn_vs_resnet}}
\end{figure}

@@

While it is not easy to explain these differences in accuracy, we hypothesize this being related to the nature of peculiar spectral and temporal morphology of each instrument. For example, flute notes tend to exhibit a steep spectral rolloff, with most of the energy captured by the first few partials. Moreover, the noisy nature of the transient portions of these notes is not well represented by frequency-based descriptions such as Mel spectra. These two factors combined, could make the disentanglement of the flute from the analyzed combination more difficult.

@@

%Orchidea implements many advanced features that are not supported by our models. For example, it is able to apply symbolic constraints to the search, hence allowing only specific instrumental combinations or playing styles. It is also able to reduce the search space by applying harmonic analysis on the target sound, and create sparse solutions that do not use every instrument specified.

In order to have a fairer comparison, we did not allow Orchidea to use any of its advanced features: we did not apply any symbolic constraints or harmonic analysis and we forced it forced to use all 10 instruments in each solution. %This creates a fairer comparison, since our models are unable to create constrained or sparse solutions.
